Fractional Types

### 2013-02-21 frac.tex
a little in sec 5

vector spaces (more refined because it doesn't collpase 1/b to b; also each
vector space must have a zero vector; so we have an explicit notion of
``error'' which we can use and track at run time. Adding this error
eliminates at RUN TIME relations which are not isomorphisms I think.

a potentionally better and more semantics is based on groupoids; mention that
we can have sets (with stucture) who cardinality is fractionals and that this
would formalize nicely the negative information

@@

>>>>

The interpretation of {{PiTF}} in the category of sets and relations is
adequate in the sense that it produces a language in which every program is a
reversible relation and in which relations are first-class values. This
relational model is however, unsatisfactory, for several reasons:

\begin{itemize}
\item the type {{1/b}} is interpreted in the same way as {{b}} which gives no
  insight into the ``true meaning'' of fractionals;
\item the interpretation is inconsistent with the view of types as algebraic
  structures; for example, {{Pi}} with the empty type, is the
  categorificationi of a semiring but although fractional types in {{PiTF}}
  syntactically ``look like'' rational numbers, there is no such formal
  connection to the rational numbers;
\item finally, we have lost some delicate structure moving from {{Pi}} to
  {{PiTF}} as we can express arbitrary relations between types and not just
  isomorphisms.
\end{itemize}

For these reasons, it is interesting to consider other possible semantic
interpretations of fractionals. A natural alternative is to use the
``canonical'' compact closed category, that is finite dimensional vector
spaces and linear maps. We outline the main features of such an
interpretation:

\begin{itemize}
\item before we talk about vector spaces, we must fix an underlying
  field. For simplicity in the following discussion we choose the underlying
  field to be the field of booleans with two elements labeled \texttt{F} and
  \texttt{T}
\item each type {{b}} in {{PiTF}} is interpreted as a finite dimensional
  vector space $\mathbf{V}_b$ over the field of booleans. Of particular
  interest:
  \begin{itemize}
  \item every vector space contains a zero vector which means that the type
    {{0}} (if included in {{PiTF}} would not be the ``empty'' type --- it
    would contain the empty vector. Furthermore all combinators would be
    partial as they would have to map the zero vector to itself. 
  \item the type {{1}} is interpreted as a 1-dimensional vector space and hence
    is isomorphic to the underlying field.
  \item the fractional type {{1/b}} is interpreted as the dual vector space
    of the vector space $\mathbf{V}_b$ representing {{b}} consisting of all
    the \emph{linear functionals} on $\mathbf{V}_b$.
  \item one can then validate certain desirable properties: {{1/(1/b)}} is
    isomorphic to {{b}}; and {{eps*_b}} corresponds to a bilinear form which
    maps a dual vector and a vector to a field element.
  \end{itemize}
\end{itemize}

Although this semantics appears to have ``better'' properties than the
relational one, we argue that it is not \emph{the} ``perfect'' semantics. By
including the zero vector, the language has again additional partial
morphisms that do not correspond to isomorphisms and it is difficult to
reconcile the interpretation with the view that the types are rational
numbers. What would really be a ``perfect'' interpretation is one in which we
can only express {{Pi}}-isomorphisms as first class values and in which the
types are interpreted in a way that is consistent with the rational
numbers. Fortunately, there is promising significant work on the groupoid
interpretation of type theory and on the categorification of the rational
numbers that we outline next.

The fundamental idea of both lines of previous work is that it is possible to
build sets (with additional structure) with a \emph{fractional cardinality}.

%% EXPAND on this

@@

vector spaces (more refined because it doesn't collapse 1/b to b; also each
vector space must have a zero vector; so we have an explicit notion of
``error'' which we can use and track at run time. Adding this error
eliminates at RUN TIME relations which are not isomorphisms I think.

a potentially better and more semantics is based on groupoids; mention that
we can have sets (with structure) who cardinality is fractionals and that this
would formalize nicely the negative information
